Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra

### 2015-05-28 ms.tex
Change "previous" and "earlier" study to Paper I
From acb28a1:
============

In ms.tex,

(*) Change "previous" and "earlier" study to Paper I
(*) Minor changes in Abstract

We confirm the recently predicted ``pitchfork'' foreground signature in power spectra of high-redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherent response of a wide-field instrument and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. In our recent paper, many signatures in the simulation which predicted this feature were validated against Murchison Widefield Array data but this key element was at or below the noise level. In this paper, we improve the data sensitivity through coherent averaging of 12 independent snapshots with identical instrument settings, and provide the first confirmation of the prediction with a signal-noise ratio $>10$. This wide-field effect can by mitigated by careful antenna design which suppresses sensitivity towards the horizon. With future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array, based on simple scaling relations, we estimate a suppression of $\gtrsim 40$~dB is required near the horizon for signal detection in cosmological H~{\sc i} modes in analyses that deploy the {\it foreground avoidance} technique.

@@

In the single snapshot (top), similar to the one used in our earlier study, faint features associated with the {\it pitchfork} signature are visible near the horizon limits. But the high level of thermal fluctuations makes their significance marginal. In contrast, the dynamic range in the averaged data (middle) is a factor $\gtrsim 10$ higher (in delay power spectrum) relative to that in a single snapshot, and is consistent with the improvement expected from averaging 12 independent snapshots. Hence, the foreground power near the horizon limits appears $\gtrsim 10$ times more prominent. Also, faint horizontal features, not seen in the single snapshot, appear at $\tau \simeq \pm 0.78\,\mu$s thus confirming the improvement in sensitivity. We identify these faint features as the response of the MWA coarse band edges flagged periodically every 1.28~MHz.

@@

In our earlier study, we proposed a selective flagging of data on different baselines that can potentially mitigate foreground contamination by two orders of magnitude. Following the confirmation presented here, efforts are underway to incorporate this proposed foreground mitigation technique into the MWA data analysis.

For future work, we plan to extend our analysis to HERA. Based on our earlier study, such a dish will have a much desirable Fourier response from a foreground contamination viewpoint. One of our objectives is to forecast the per-baseline foreground contamination as a function of local sidereal time in order to tune the HERA observing strategy and data analysis to maximize sensitivity to the EoR signal.

>>>>

We confirm the recently predicted ``pitchfork'' foreground signature in power spectra of high-redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherent response of a wide-field instrument and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. In our recent paper, many signatures in the simulation which predicted this feature were validated against Murchison Widefield Array data but this key element was at or below the noise level. In this paper, we improve the data sensitivity through coherent averaging of 12 independent snapshots with identical instrument settings, and provide the first confirmation of the prediction with a signal-noise ratio $>10$. This wide-field effect can by mitigated by careful antenna design which suppresses sensitivity towards the horizon. With future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array, based on simple scaling relations, we estimate a suppression of $\gtrsim 40$~dB is required near the horizon for signal detection in cosmological H~{\sc i} modes in analyses that employ {\it foreground avoidance} techniques.

@@

In the single snapshot (top), similar to the one used in Paper~I, faint features associated with the {\it pitchfork} signature are visible near the horizon limits. But the high level of thermal fluctuations makes their significance marginal. In contrast, the dynamic range in the averaged data (middle) is a factor $\gtrsim 10$ higher (in delay power spectrum) relative to that in a single snapshot, and is consistent with the improvement expected from averaging 12 independent snapshots. Hence, the foreground power near the horizon limits appears $\gtrsim 10$ times more prominent. Also, faint horizontal features, not seen in the single snapshot, appear at $\tau \simeq \pm 0.78\,\mu$s thus confirming the improvement in sensitivity. We identify these faint features as the response of the MWA coarse band edges flagged periodically every 1.28~MHz.

@@

In Paper~I, we proposed a selective flagging of data on different baselines that can potentially mitigate foreground contamination by two orders of magnitude. Following the confirmation presented here, efforts are underway to incorporate this proposed foreground mitigation technique into the MWA data analysis.

For future work, we plan to extend our analysis to HERA. Based on Paper~I, such a dish will have a much desirable Fourier response from a foreground contamination viewpoint. One of our objectives is to forecast the per-baseline foreground contamination as a function of local sidereal time in order to tune the HERA observing strategy and data analysis to maximize sensitivity to the EoR signal.
